---
title: "data_prep_UFC"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, error = FALSE, message=FALSE)
```

# Cleaning data from https://www.kaggle.com/datasets/rajeevw/ufcdata?select=raw_fighter_details.csv
```{r}
library(ggplot2)
library(tidyverse)
library(readr)
library(dplyr)
library(knitr)
library(kableExtra)
```
```{r}
my_knit <- function(df, n = 5) {
  kableExtra::kable_styling(knitr::kable(head(df ,n)), bootstrap_options = "striped", full_width = FALSE)
}
```

```{r importing data}
raw_fighter_details <- read_csv("raw_fighter_details.csv")
my_knit(raw_fighter_details)
```
### Removing rows with NA
```{r}
raw_fighter_details <- raw_fighter_details[!is.na(raw_fighter_details$Reach),]
raw_fighter_details <- raw_fighter_details[!is.na(raw_fighter_details$Height),]

## Replacing a NA stance with "Orthodox"

which(is.na(raw_fighter_details$Stance))

to_Orthodox <- which(is.na(raw_fighter_details$Stance))

raw_fighter_details$Stance[to_Orthodox] <- "Orthodox"

```

### Columns with % sufix "Str_Acc %" "Str_Def %" "TD_Acc %"  "TD_Def %"
```{r}
columns_percentage_suffix <- colnames(raw_fighter_details)[c(8,10,12,13)]
columns_percentage_suffix

#Replacment % with space
raw_fighter_details[,columns_percentage_suffix] <- lapply(raw_fighter_details[,columns_percentage_suffix], gsub, pattern = "%", replacement = "")

#Chenging type of variables
raw_fighter_details[,columns_percentage_suffix] <- lapply(raw_fighter_details[,columns_percentage_suffix], as.numeric)

#Adding % to the Column names
colnames(raw_fighter_details)[c(8,10,12,13)] <- paste0(colnames(raw_fighter_details)[c(8,10,12,13)], "_%")

colnames(raw_fighter_details)[3] <- "Weight_(lbs)"

raw_fighter_details[,3] <- lapply(raw_fighter_details[,3], gsub, pattern = "lbs.", replacement = "")
raw_fighter_details[,3] <- lapply(raw_fighter_details[,3], as.numeric)
```
## Problem with the weight
#### In UFC are 9 weight clasess, in ourdata we have got 38
```{r}
length(unique(raw_fighter_details$`Weight_(lbs)`))

sort(unique(raw_fighter_details$`Weight_(lbs)`))
weights <- sort(unique(raw_fighter_details$`Weight_(lbs)`))

```
# I decided to replace wrong values with the value above, because when we assign person to the weight class, if it pass certain value
# it will be assigned to the weight class above. eg. We've got classes 205, and 265. 
# All persons with weight above 205 are in 265 weight class
```{r}
weights[weights > 205]
weights[weights > 205 & weights <= 265]

to_265 = weights[weights > 205]

to_145 = c(139, 140)

to_170 = 168

raw_fighter_details["Weight_(lbs)"] <- sapply(raw_fighter_details["Weight_(lbs)"],
                             function(x) replace(x, x %in% to_265, 265))

raw_fighter_details["Weight_(lbs)"] <- sapply(raw_fighter_details["Weight_(lbs)"],
                             function(x) replace(x, x %in% to_145, 145))

raw_fighter_details["Weight_(lbs)"] <- sapply(raw_fighter_details["Weight_(lbs)"],
                             function(x) replace(x, x %in% to_170, 170))

sort(unique(raw_fighter_details$`Weight_(lbs)`))

```

# Prepering dataframe for EU standard
```{r}
# Separating a weight column, to make kg cloumn 

weight <- as.data.frame(raw_fighter_details$`Weight_(lbs)`)

colnames(weight) <- "Weight_(kg)"

weight <- weight %>% 
  transmute(`Weight_(kg)` = round(`Weight_(kg)`*0.45359237))

## Reach

colnames(raw_fighter_details)[4] <- "Reach_(inches)"

raw_fighter_details[,4] <- lapply(raw_fighter_details[,4], gsub, pattern = '"', replacement = "")
raw_fighter_details[,4] <- lapply(raw_fighter_details[,4], as.numeric)

# Making a reach column in centimeters
reach <- as.data.frame(raw_fighter_details$`Reach_(inches)`)

colnames(reach) <- "Reach_(cm)"

reach <- reach %>% 
  transmute(`Reach_(cm)` = round(`Reach_(cm)`*2.54))

## Height
# 1feet = 12 inches

height <- lapply(raw_fighter_details[,2], str_split, pattern = " ", simplify = TRUE)

height <- as.data.frame(height)
colnames(height) <- c("Foots","Inches")


height[,1] <- vapply(height[,1], gsub, pattern = "\'", replacement = "", FUN.VALUE = character(1))
height[,2] <- vapply(height[,2], gsub, pattern = "\"", replacement = "", FUN.VALUE = character(1))

height[,c(1:2)] <- lapply(height[,c(1:2)], as.numeric)

height <- height %>%
  mutate(Inches_only = Foots*12 + Inches,
         cm = round(Inches_only*2.54))


## Making EU and USA unit based dataframe

## USA
raw_fighter_details_USA <- raw_fighter_details

## EU
raw_fighter_details_EU <- raw_fighter_details %>%
  mutate("Height" = height$cm,
         "Weight_(lbs)" = weight$`Weight_(kg)`,
         "Reach_(inches)" = reach$`Reach_(cm)`)

colnames(raw_fighter_details_EU)[2:4] <- c("Height_(kg)", "Weight_(kg)", "Reach_(cm)")
```

```{r}
my_knit(raw_fighter_details_EU)
my_knit(raw_fighter_details_USA)

```

